---
title: "Short Homework 2"
author: "Bhavika Sewpal - 300089940"
date: "4/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\
**Question 1**
```{r}
set.seed(100)
X1 = rnorm(100,sd=2)
X2 = rnorm(100,sd=2)
epsilon = rnorm(100,sd=1)
Y = 1.8 + 0.6 * X1 + 1.2 * X2 + epsilon

#Choose beta_1 = 0.6 
a = Y - 0.6 * X1
fit1 = lm(a~X2)
fit1$coefficients[2]

#choose beta_2
a = Y = 1.2 * X2
fit2 = lm(a~X1)
fit2$coefficients[2]

beta_0 = rep(0,1000)
beta_1 = rep(0,1000)
beta_2 = rep(0,1000)

#set the first value of beta_1 to 5
beta_1[1] = 5

for (i in 1:1000){
  a = Y - beta_1[i] * X1
  beta_2[i] = lm(a~X2)$coefficients[2]
  a = Y - beta_2[i] * X2
  fit = lm(a~X1)
  if(i < 1000){
    beta_1[i+1] = fit$coefficients[2]
  }
  beta_0[i] = fit$coefficients[1]
}
plot(1:1000,beta_0,type="l",xlab="Iterations",ylab="betas",col="red",ylim=c(-0.02,2))
lines(1:1000,beta_1,col="blue")
lines(1:1000,beta_2,col="green")
legend("center",c("beta_0", "beta_1", "beta_2"), lty = 1, col = c("red", "blue", 
    "green"))
```
\
**Note**:\
The lines for $\beta_0$ and $\beta_1$ overlap. 


```{r}
fit2 = lm(Y ~ X1 + X2)
plot(1:1000, beta_0, type = "l", xlab = "Iterations", ylab = "betas", ylim = c(-0.02, 
    2), col = "red")
lines(1:1000, beta_1, col = "blue")
lines(1:1000, beta_2, col = "green")
abline(h = fit2$coef[1], lty = "dashed", lwd = 3)
abline(h = fit2$coef[2], lty = "dashed", lwd = 3)
abline(h = fit2$coef[3], lty = "dashed", lwd = 3)
legend("center", c("beta0", "beta1", "beta2", "mlr"), lty = c(1, 
    1, 1, 2), col = c("red", "blue", "green", "black"))


```
\
The multiple regression coefficients match the backfitting coefficients, as illustrated in the above diagram.\
In this case, 1 backfitting iteration was sufficient to obtain the same results as in multiple regression.\
**Question2**\

```{r}
#Creating dataset

set.seed(200)
x1 <- rnorm(1000, 0, 1)          
y1 <- 5*x1^2 + runif(1000, 1,1.2) 

x2<- rnorm(1000, 0, 1)          
y2 <- 5*x2^2 + runif(1000, 20, 20.2) 

x <- c(x1,x2,y1,y2)
x <- matrix(x,ncol=2)
y <- c(rep(1,1000),rep(2,1000))
plot(x,col=y)

y <- c(rep(1,1000),rep(2,1000))
df_x <- data.frame(c(x1,x2),c(y1,y2))
df <- data.frame(x=df_x,y=y)
df[,3] <- factor(df[,3])
plot(df[,1],df[,2])

train_index <- sample(c(1:2000),1000)
train <- df[train_index,]
test<- df[-train_index,]
library (e1071)
 #Linear svm
tune.out=tune(svm ,y~.,data=train ,kernel ="linear",ranges =list(cost=c(0.001 , 0.01, 0.1, 1,5,10,20,30,40,50,100) ))
#c(0.001 , 0.01, 0.1, 1,5,10,50,100) )
bestmod =tune.out$best.model
summary (bestmod )
plot(bestmod,train)
train_pred <- predict(bestmod,train)
truth <- train$y
table(train_pred,truth)

test_pred <- predict(bestmod,test)
truth <- test$y
table(test_pred,truth)

#polynomial svm
tune.out=tune(svm ,y~.,data=train ,kernel ="polynomial",ranges =list(cost=c(0.001 , 0.01, 0.1, 1,5,10,20,30,40,50,100) ), degree=2)
#c(0.001 , 0.01, 0.1, 1,5,10,50,100) )
bestmod =tune.out$best.model
summary (bestmod )
plot(bestmod,train)
train_pred <- predict(bestmod,train)
truth <- train$y
table(train_pred,truth)

test_pred <- predict(bestmod,test)
truth <- test$y
table(test_pred,truth)
```
\
The trainig and testing errors are greater than 15% in the the case of a linear svm.\

**Question3**\
```{r}
library(RSSL)
data<-generateCrescentMoon(150,2,1)
plot(data$X1,data$X2,col=data$Class,asp=1)
library(stats)
km <- kmeans(data[,2:3],2,nstart=20)
summary(km)
```
\
With this dataset, the algorithm gets stuck on a local maxima.

**Question4**\
a) \
There is not enough information to tell.\

Single linkage uses the smallest distance between the elements, whereas complete linkage uses the largest.\
If it happens that these distances are equal, then fusion will occur at the same height. \
If these distances are not equal, the height of fusion for complete linkage will be higher than the height of fusion for single linkage.\

b)\
Fusion will occur at the same height for single elements.


**Question5**\
